---
output: html_document
---
<!--
%\VignetteEngine{knitr::docco_linear}
%\VignetteIndexEntry{Introduction to reserve}
-->

```{r, echo = FALSE, message = FALSE}
library(reserve)
knitr::opts_chunk$set(
  comment = "#>",
  tidy = FALSE)
```

# Introduction to reserve
The goal of this document is to introduce the basic tools `reserve` provides, and to walk through a realistic example using the `reserve` package.  This vignette and package are nowhere near complete.  I am creating the vignette in conjunction with the package to help keep my thoughts in line, and to remind me where I left off.

The following Background section is taken from the `ChainLadder` package vignette [Claims reserving with R: Chainladder-0.1.7 Package Vignette](http://cran.r-project.org/web/packages/ChainLadder/vignettes/ChainLadder.pdf) by Markus Gesmann.

## Background: Claims reserving in Insurance

Unlike other industries the insurance industry does not sell products as such, but promises.  An insurance policy is a promise by the insurer to the policyholder to pay for future claims for an upfront received premium.

As a result insurers don't know the upfront cost of their service, but rely on historical data analysis and judgement to derive a sustainable price for their offering.  In General Insurance (or Non-Life Insurance, e.g. motor, property and casualty insurance) most policies run for a period of 12 months.  However, the claims payment process can take years or even decades.  Therefore often not even the delivery date of their product is known to insurers.  

In particular claims arising from casualty insurance can take a long time to settle.  Claims can take years to materialise.  A complex and costly example are the claims from asbestos liabilities.  A research report by a working party of the Institute of Actuaries has estimated that the undiscounted cost of UK mesothelioma-related claims to the UK Insurance Market for the period 2009 to 2050 could be around 10bn pounds.  The cost for asbestos related claims in the US for the worldwide insurance industry was estimated to be around $120bn in 2002.

Thus, it should come to no surprise that the biggest item on the liability side of an insurer's balance sheet is often the provision or reserves for future claims payments.  Those reserves can be broken down in case reserves (or out-standing claims), which are losses already reported to the insurance company and incurred but not reported (IBNR) claims.

## Motivation

It is often difficult to create reproducible actuarial reserve reports using a spreadsheet gui.  R has many advantages over spreadsheets (in R you type commands rather than pointing and clicking, R can work with larger datasets, etc.). The goal of the `reserve` package is to provide a standard process for creating reserve reports using R. `reserve` provides a standard means of storing your claims data so you can use the same set of functions (defined by `reserve`) to automate many repetetive reserving tasks and create reproducible reports that are generated directly from the original claim loss detail.

When a reserve report is not directly created from the the original loss detail the report loses flexibility (e.g. If you create a development triangle using no excess retention limits it can be very time time consuming to create a new development triangle based on an excess retention limit of say 250,000 because the data is summarized on an origin period basis; claim detail is lost). This practice of using data summarized on an origin period basis is popular when using spreadsheets because keeping all the historical loss information in one spreadsheet quickly results in data sets too large for the spreadsheet to handle.  I run into problems after a couple hundred thousand rows using Excel 2010.

The `reserve` package improves flexibility in reserve report creation by maintaining access to all the original claim detail (i.e. each report is generated from the original claim data as opposed to a summary of the original claim data); `reserve` makes the reserve report generation process reproducible.  `reserve` provides a standard format for storing claim loss detail and provides a set of functions specific to that standard format.  

## Getting Started

The standard format `reserve` uses to store claim loss data is the S3 class `loss_df`.  A `loss_df` is created using the `loss_df()` function.  A `loss_df` holds and cetegorizes loss information on a claim or occurence basis, and is flexible enough to handle data that is categorized and organized differently from company to company. 

The rest of this vignette will guide you through an example using reserve.  The example will proceed as follows:
* Taking a data set (`losses`) in a format resembling how data may be provided by an insurance company, and transforming that data set into a `loss_df` object that contains and organizes all the information relevant to the reserve analysis.
* Reviewing the `loss_df` for errors and probelm areas.
* Performing statistical reserving techniques on the `loss_df`.
* Create a PDF document of the results complete with graphics, tables, and text.

View the structure of the uncleaned `losses` data frame using the `str` function:
```{r, eval = FALSE}
str(losses)
```
There are many columns that are not relevant to reserving analysis, and some transformations must be applied to the data to get all the necessary information out of the losses data frame.  Let's get started.

### Format the data 

The `loss_df()` function takes the uncleaned data frame as its first formal argument.  All additional arguments are used to categorize the columns in the loss df.  There are two main categories that a loss_df column can fall into: required (general information about the claim) and loss values (loss dollar amounts and claim descritions)

* Required: These categories provide all the necessary claim detail other than the actual loss values. A `loss_df` is required to have the following columns:
  * id: factor - a unique identification key for each claim
  * origin: numeric - the time period in which the claim originated
  * dev: numeric - the development stage of the claim at the relevant evaluation_date

* Loss Values: All loss values must be numeric. Each category can contain as many columns as necessary, but avoid double counting (i.e. do not have one category for paid losses, one for paid expenses, and one for total paid losses and expenses.  Just supply the paid losses and paid expenses categories separately or the sum of paid losses and paid expenses).  All loss values must fall into 1 of the following 5 categories:
  * paid: paid amounts
  * incurred: incurred amounts
  * paid_recovery: paid recovery amounts
  * incurred_recovery: incurred recovery amounts
  * desc: miscellaneous descriptive amounts

Next the dplyr package is used to clean the `losses` data so it conforms to the format required to make a `loss_df`.  If you are not familiar with the dplyr package see the [dplyr introduction vignette](http://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html)
```{r, message = FALSE}
library(dplyr)
# create origin and dev column
losses <- mutate(losses, origin = as.numeric(substr(fiscal_year_desc, 1, 4)), 
                 evaluation_year = as.numeric(format(as.Date(evaluation_date, "%Y-%m-%d"), "%Y")),
                 dev = evaluation_year - origin) 

# group claims by occurence
# often necessary when excess reinsurance is applied on an occurence basis
# rather than on a claims basis.
occurences <- losses %>%
  group_by(claim_number, dev, origin) %>%
     summarise(claim_cts = n(),
               payment_amount = sum(payment_amount), # paid loss & ALAE
               reserve_amount = sum(reserve_amount), # incurred loss & ALAE
               paid_expense = sum(X4_exp_payment), # paid ALAE
               incurred_expense = sum(X4_exp_reserve), #incurred ALAE
               sal_sub = sum(payment_no_reserve_a), # salvage and subrogation
               sal_sub_incurred = sum(payment_no_reserve_a)
              )

# create relevant loss values columns
occurences <- mutate(occurences,
                    paid_loss_only = payment_amount - paid_expense,
                    incurred_loss_only = reserve_amount - incurred_expense,
                    paid_excess250 = max(payment_amount - sal_sub - 250000, 0),
                    incurred_excess250 = max(reserve_amount - sal_sub - 250000, 0)) 
```
The above code does not use the `reserve` package.  All that code did was prepare the data for use with the `reserve` package (often a time consuming task the first time around).  You can use any of the tools provided by R for data preparation; you do not need to use dplyr.

Now that we have all the necessary columns, we can create the `loss_df` object.
```{r}
# create loss_df object
mydf <- loss_df(occurences, id = "claim_number",
                             origin = "origin",
                             dev = "dev", 
                             paid = c("paid_loss_only", "paid_expense"),
                             incurred = c("incurred_loss_only", "incurred_expense"),
                             paid_recovery = c("paid_excess250", "sal_sub"),
                             incurred_recovery = c("incurred_excess250", 
                                                   "sal_sub_incurred"),
                             desc = "claim_cts"
                 )
head(mydf[, 1:6])
```

Each column has an attribute specifying the `type` of loss detail that the column contains.  The `type` was defined by the argument the column was supplied to in the `loss_df()` function (i.e. `paid_loss_only` and `paid_expense` are of `type` `paid`).

### Review the data

Now we can use the `reserve` package to review the data.  Let's start by seeing a summary of the most recent loss detail summarized by origin period.
```{r}
summary(mydf)
```

We can look at the data at an older evaluation date by specifying the `evaluation_date`
```{r}
summary(mydf, evaluation_date = "2012")

```

and for a bar chart of the data...
```{r}
plot(mydf)
```

and at an alternative `evaluation_date`
```{r}
plot(mydf, evaluation_date = "2012")
```

We can return a data frame of all the claims that have experienced a change from one evaluation date to another by using the `claim_changes()` function:

```{r}
# specify the loss amount values you want to see the changed claims for 
mychanges <- claim_changes(mydf, eval1 = "2013", eval2 = "2012",
                            values = c("paid_loss_only", "claim_cts"))
head(mychanges)
```

`mychanges` is a data frame consisting of all the claims in which there was a change in the "paid_loss_only" column from 2012-06-30 to 2013-06-30.  You can now browse through the changed claims to spot obvious problems with the new data.  For example we may want to check that there are no missing claims (i.e. no claims that were in the data at the last `evaluation_date` that are no longer in the data)

```{r}
# check for missing claims
mychanges[, mychanges$claim_cts_change < 0]
```
We may also want to check if the `paid_loss_only` category decreased for any claims.
```{r}
# check for claims in which paid_loss_only decreased
mychanges[, mychanges$paid_loss_only_changes < 0]
```

There are no missing claims or claims with a decrease in `paid_loss_only`.  After this quick review it appears that new data are reasonable.  Next we can project some ultimate losses. 

### Create Projections

Before a projection is made we must specifiy the loss amounts we wish to project (e.g. paid loss & ALAE gross of all recoveries, paid loss & ALAE net of all recoveries, medical only paid loss & ALAE gross of all recoveries, etc.).  The `retain` function specifies the values to be projected.
```{r}
# include recovery detail by category: `sal_sub` and `paid_excess250`
value2project <- retain(mydf, value = "paid", recovery = c("paid_excess250", "sal_sub"))
head(value2project)
```

Now the `ChainLadder` package can be used to make projections.  