---
output: 
  html_document:
    theme: flatly
---
<!--
%\VignetteIndexEntry{Introduction to lossdb}
%\VignetteEngine{knitr::rmarkdown}
-->

```{r, echo = FALSE, message = FALSE}
library(lossdb)
library(knitr)
knitr::opts_chunk$set(
  comment = "#>",
  tidy = FALSE)
```

# Introduction to lossdb

Warning: This vignette and package are still in the early stages of development...

## Background: Actuaries and Insurance Loss Data

Actuaries often work with data sets describing the dollar value and nature of insurance losses (loss data).  It is common for actuaries to analyze loss data on a "by id" basis (i.e. each row/observation represents a claim/occurrence/member at a certain development stage) or on  a "by origin" basis (i.e. each row/observation represents a policy/accident period at a certain development stage).  If data is not provided by id or by origin it is transformed into one of the two formats for analysis.

## Motivation

I want to make reproducible actuarial reports in R.  Getting raw data ready for analysis is a pain, and once the data is ready it is nice to have it in some kind of a standard format so you can do similar things to it whenever you get it into that format... enter `lossdb`

The goal of the `lossdb` package is to provide standard functions for manipulating, visualizing, and modeling loss data. `lossdb` provides a standard means of storing your loss data on a by id or by origin basis so you can use the same set of functions (defined by `lossdb`) to automate many repetitive tasks and ultimately create reproducible reports that are generated directly from the original loss data.  

## Philosophy

Columns in loss data can be losely grouped into a few different categories.  By organizing each column into one of these categories (defined below) certain actuarial analysis tasks can be automated across many loss data sets. 

### The Structure of Loss Data

`lossdb` groups all the loss data columns applicable to the analysis into one of 2 main categories.  These two overarching categories are further subdivided into several more categories.  A description of the `lossdb` stucture is detailed below:

* meta
    + id
    + dev
    + origin
* detail
    + dollar
        + paid
        + incurred
        + paid_recovery
        + incurred_recovery
    + desc

More detail on the lossdb structure is described below:

#### meta

The 3 meta columns are of particular importance.  They are defined as follows:

* origin - required - numeric - single column - the period in which the loss is associated with
* dev - required - numeric - single column - the development stage of the loss
* id - optional - factor - single column - an identifier for the loss (it should be unique by a dev and origin period)

#### dollar

All columns representing dollar amounts are "dollar" columns.  `lossdb` assumes that all the information contained in dollar columns can be sub categorized into 4 different groups without losing any information.  Each of these 4 groups can contain as many columns as neccessary.  All dollar groups are optional, but you need to provide at least one column for one of the groups.  The columns must be numeric and you can have as many columns in each group as desired.  The 4 groups:

* paid - actual amounts already paid
* incurred - amounts incurred (i.e. total paid + reserve booked to be paid by the insurance company before the claim is closed)
* paid_recovery - actual paid reimbursements
* incurred_recovery - incurred reimbursements

#### desc

Other columns included in loss data provide some type of description of the claim (e.g. claimant name, whether the claim is open or closed, etc.).  I refer to all of these columns as "desc" (short for description) columns.  The `lossdb` package can handle any number of description columns, they are optional, and they can be of any type.

# Getting Started - An Example

The example will proceed as follows:

1. Take a data set (`occurrences`) and transform that data set into a `loss_df` object that contains and organizes all the information relevant to the analysis.
2. Review the `loss_df` for errors and potential problem areas.
3. Perform statistical reserving techniques on the `loss_df` using the `ChainLadder` package.

View the structure of the `occurrences` data frame using the `str` function:
```{r}
str(occurrences)
```

Now we can create the `loss_df` object.

```{r}
# create loss_df object
mydf <- loss_df(occurrences, 
          id = "claim_number",
          origin = "origin",
          dev = "dev", 
          paid = c("paid_loss_only", "paid_expense"),
          incurred = c("incurred_loss_only", "incurred_expense"),
          paid_recovery = c("paid_excess250", "sal_sub"),
          incurred_recovery = c("incurred_excess250", 
                                "sal_sub_incurred"),
          desc = "claim_cts"
        )
kable(head(mydf[, 1:6]))
```

Each detail (dollar or desc) column has an attribute specifying the type of loss detail that the column contains.  This attribute is named the "detail" attribute.  The detail attribute of each column is defined by the argument the column is supplied to in the `loss_df()` function (i.e. `paid_loss` and `paid_expense` have a detail attribute of "paid").  All detail columns maintain the column names that they are supplied with.  The names for `meta` columns are changed to the `meta` category to which they were supplied.

### Review the data

Now we can use the `lossdb` package to review the data.  Let's start by seeing a summary of the most recent `calendar` period (`calendar` = `origin` + `dev`) summarized by `origin` period.

```{r summary1}
kable(summary(mydf)[, 1:9])
```

We can look at the data at an older `calendar` period by specifying the `calendar` argument in the `summary()` function.

```{r summary2}
kable(summary(mydf, calendar = "2012")[, 1:9])
```

Note: the `calendar` period is the `origin` period plus the `dev`. (e.g. The `calendar` for all claims in origin year 2010 at their first `calendar` period would be 2011.)

and the built in bar chart representation of the data...

```{r}
plot(mydf)
```

and plotted at an alternative `calendar`

```{r}
plot(mydf, calendar = "2012")
```

We can return a data frame of all the claims that have experienced a change from one calendar to another by using the `claim_changes()` function:

```{r}
# specify the loss amount values you want to see the changed claims for 
mychanges <- claim_changes(mydf, 
               calendar1 = "2013", 
               calendar2 = "2012",
               values = c("paid_loss_only", "claim_cts")
             )
kable(head(mychanges))
```

`mychanges` is a data frame consisting of all the claims in which there was a change in the `paid_loss` or `claim_cts` column from calendar period 2012 to 2013.  You can now browse through the changed claims to spot obvious problems with the new data.  For example we may want to check that there are no missing claims (i.e. no claims that were in the data at the last `calendar` that are no longer in the data)

```{r}
# check for missing claims
kable(mychanges[mychanges$claim_cts_change < 0, ])
```

This check revealed that there are no missing claims in our `loss_db` from calendar 2012 to 2013.

We may also want to check if the `paid_loss` category decreased for any claims.

```{r}
# check for claims in which paid_loss decreased
kable(mychanges[mychanges$paid_loss_only_change < 0, ])
```

There are a few claims with a decrease in "paid_loss_only".  Claims should not decrease in gross paid loss as they develop, but it happens in real world loss data.  Fortunately none of the paid amounts decreased so significantly that we need to stop our analysis and investigate.  Next we can project some ultimate losses. 

### Create Projections

Before a projection is made we must specifiy the loss amounts we wish to project (e.g. paid loss & ALAE gross of all recoveries, paid loss & ALAE net of all recoveries, medical only paid loss & ALAE gross of all recoveries, etc.).  Use the `paid()`, `incurred()`, `paid_recovery()`, and `incurred_recovery()` functions to get the total from each respective "dollar" category.

```{r projection_values}
# project total paid losses gross of any recovery
value2project <- data.frame(origin = mydf$origin, 
                   dev = mydf$dev, 
                   paid_total = paid(mydf)
                 )
kable(head(value2project))
```

Now the `ChainLadder` package can be used to make projections.  

```{r triangle, message = FALSE}
library(ChainLadder)
paid_tri <- as.triangle(value2project, 
              origin = "origin", 
              dev = "dev", 
              value = "paid_total"
            )
```

```{r mack, warning = FALSE}
MackChainLadder(paid_tri)
```

```{r boot, warning = FALSE}
BootChainLadder(paid_tri)
```